### 2023-12-04 scripts/Regress_AIM_cover_Patch.Rmd
adding empty scripts for steps

---
title: "Regress Percent Cover of Species in AIM plots against Mean Percent Patch Suitability"
author: "steppe"
date: "2023-12-04"
output: html_document
---

Species Distribution Models have been shown to be semi-quantitative, i.e. be able to predict species abundance from suitability scores,  under various scenarios.
Here we determine whether a positive relationship exists between the mean predicted habitat suitability of a patch, or a sector of large patches, and the percent aerial cover of the taxon on an AIM plot.