### 2023-12-18 scripts/generate_absences.Rmd
problems with random sampler, aim sample size resolved

---
title: "Generate Real and Pseudo-absences"
author: "steppe"
date: "2023-12-06"
output: html_document
---

```{r load packages}
library(tidyverse)
library(sf)
library(terra)
source('functions.R')
```

```{r Create simplified BLM surface management, eval = F}
p1 <- '/media/steppe/hdd/geospatial/SMA_WM.gdb'

st_layers(p1)
ag <- st_read(p1, 'SurfaceMgtAgy_BLM') %>% 
  filter(ADMIN_ST != 'AK') %>% 
  rename(geometry = SHAPE) %>% 
  select(geometry) %>% 
  st_make_valid() %>% 
  st_union()

ag <- st_simplify(ag, dTolerance = 1000)
plot(ag)

pout <- '/media/steppe/hdd/SDM_restorations/data/raw/coarse_blm/'
st_write(ag, dsn = paste0(pout, 'coarse_surface.shp'))
```

We use absences while creating species distribution models. 
Both true absences, acquired from vegetation monitoring plots, and pseudo absences generated via geographic and environmental space are used.

```{r load aim data for absences, eval = F}

p  <- '../data/raw/2021_AIM_Terrestrial/AIMTerrestrial9-1-22.gdb'
AIM_points <- st_read(dsn = p, layer = 'TerrADat', quiet = T) %>%
  select(PrimaryKey, DateVisited, geometry = Shape) %>% 
  st_transform(5070)

pout <- '/media/steppe/hdd/SDM_restorations/data/raw/coarse_blm/'
blm_surf <- st_read(paste0(pout, 'coarse_surface.shp'), quiet = TRUE)
blm_surf <- st_simplify(blm_surf, dTolerance = 500) # just make it a bit smaller. 
blm_surf <- st_make_valid(blm_surf)

rm(p, pout)
```

After we have removed spurious occurrence records (see chunk: 'subset occurrence records for final clean batch', in clean_records), we can generate absences for each species. 
```{r generate raw absences, eval = F}

# here we subset a terrestrial surface to our area of analysis - we don't want all of
# our random points to fall in the ocean

domain <- rast(nrows = 1, ncols = 1) # create a big empty raster, you can go in through sf too. 
ext(domain) <- c( -125.5, -100, 27,  50) # set the extent
crs(domain) <- "EPSG:4326" # define the projection
domain <- as.polygons(domain) |>  # convert to vector data
  st_as_sf() |> # convert to simple feature
  st_bbox()

land <- rnaturalearth::ne_download(type= 'land', scale = 10, category= 'physical') |>
  sf::st_as_sf()
land <- st_crop(land, domain) |>
  st_transform(5070)

rm(domain)

records <- st_read('../data/raw/occurrence/thinned/thinned_occurrences.shp', quiet = TRUE)
records <- records %>% 
  group_by(taxon) %>% 
  filter(n() >= 15)
records <- split(records, f = records$taxon)

out1 <- lapply(records, absence_drawer, terrestrial = land,  bg_abs = 0.15)

out <- bind_rows(out1) %>% 
  rename(taxon = Taxon) %>% 
  mutate(date = as.Date('2023-12-18'), .before = geometry)

st_write(out, '../data/raw/occurrence/absences/absences.shp', append = FALSE)

rm(out1, blm_surf, AIM_points, records)
```

```{r extract environmental predictors to records}

absences <- st_read('../data/raw/occurrence/absences/absences.shp', quiet = TRUE)
presences <- st_read('../data/raw/occurrence/thinned/thinned_occurrences.shp', quiet = TRUE) %>% 
  mutate(Occurrence = 1, 
         PtType = if_else(is.na(PrmryKy), 'AIM Presence', 'Opportunistic Presence'), .before = geometry)

abvi_p <- filter(presences, taxon == 'Abronia villosa')
abvi_a <- filter(absences, taxon == 'Abronia villosa')

abvi <- bind_rows(abvi_p, abvi_a) %>% 
  st_as_sf() %>% 
  mutate(ID = 1:n())

table(abvi$Occurrence)

p <- '/media/steppe/hdd/Geospatial_data/WesternPlantPredictors2/chelsa'
paths <- file.path(p, list.files(p, recursive = F))
files <- lapply(paths, function(x){ file.path(x, list.files(x)) })
names(files) <- basename(paths)
bio <- lapply(files, vrt)
bio <- rast(bio)

vif_vals <- spatSample(bio, size = 5000)
collinear_probs <- vifstep(vif_vals, th = 10)
exclude()

abvi_rv <- extract(bio, abvi)
abvi_rv <- select(abvi_rv, -gdgfgd10, -gdgfgd5, -swe) %>% 
  drop_na()
abvi_rv_scaled <- data.frame(apply(abvi_rv[,2:17], MARGIN = 2, FUN = scales::rescale, to = c(0,1)))
abvi_rv_scaled$ID <- abvi_rv$ID

table(abvi_rv_scaled$Occurrence)

abvi <- right_join(abvi, abvi_rv_scaled, by = 'ID')

table(abvi$Occurrence)

rm(abvi_a, abvi_p, p, paths, files, abvi_rv)
```

```{r remove pseudo absences in presence niche space}

library(klaR)
library(psych)
library(MASS)
library(ggord)

# to determine whether the pseudo-random absences are in areas which are possibly suitable
# we will use linear discrimination analysis (LDA). LDA uses a label, e.g. presence or absence
# as a response which can be predicted by independent variables which are linear combinations 
# generated by reducing a higher dimensional data set to two axis. LDA will be trained on a data set 
# of AIM Plots with true absences, geographic absences, and a subset of the Psuedo-absences. 
# Only up to 750:750 presences:absences per species will be used to train a model. The maximum
# imbalance between presences to absences will be 750:250, or 3:1.

lda_PA_dropper <- function(x, path, column_range){

  pres <- x[x$Occurrence == 1, ]
  real_abs <- x[x$PtType == 'AIM Absence', ]
  geo_abs <- x[x$PtType == 'Background Absence', ]
  model_abs <- dplyr::bind_rows(real_abs, geo_abs)

  if(nrow(pres) > 750){pres <- pres[sample(1:nrow(pres), size = 750, replace = FALSE),]}
  if(nrow(model_abs) > 750){model_abs <- model_abs[sample(1:nrow(model_abs), size = 750, replace = FALSE),]}

  training <- dplyr::bind_rows(pres, model_abs) 
  training <- training[, c('Occurrence', colnames(abvIng)[7:22])]
  testing <- filter(x, ! ID %in% training$ID)

  LDA_model <- MASS::lda(Occurrence ~ ., training) # generate model

  # Here the accuracy of the trained modeled is evaluated
  LDA_train <- predict(LDA_model, training)$class
  tab <- table(Predicted = LDA_train, Actual = training$Occurrence)
  ACC_train <- sum(diag(tab))/sum(tab) # accuracy

  # Here the model is applied to unclassified data
  LDA_test <- predict(LDA_model, testing)$class
  testing$Predicted <- predict(LDA_model, testing)$class
  tab1 <- table(Predicted = LDA_test, Actual = testing$Occurrence)
  ACC_test <- sum(diag(tab1))/sum(tab1) # accuracy.

  n_pres <- nrow(x[x$Occurrence == 1,])
  n_abs <- nrow(x[x$Occurrence == 0,])
  conflicted <- filter(testing, Predicted != Occurrence) 
  max_rcs_2_remove <- n_abs - n_pres

  if(nrow(conflicted) == 0){ # identify the records we want to remove from the test data. 
    removals <- NA
  } else if (nrow(conflicted) > max_rcs_2_remove) {
    removals <- conflicted[sample(1:nrow(conflicted), size = max_rcs_2_remove, replace = FALSE), 'ID']
  } else {
    removals <- conflicted$ID
  }

  if(!is.na(removals)){data_out <- dplyr::filter(x, ! ID %in% removals)}

  metrics <- data.frame(
    Train_Presences = nrow(pres), 
    Train_Absences = nrow(model_abs),
    ACC_train = ACC_train, 
    ACC_test = ACC_test,
    Flagged = nrow(conflicted), 
    Removed = length(removals)
    )

  species <- x$taxon[1]
  message(length(removals), ' records dropped from ', species)
  species <- gsub(' ', '_', species)
  write.csv(metrics, file.path(path, paste0(species, '.csv')), row.names = FALSE)

  return(data_out)
}

abvIng <- st_drop_geometry(abvi) %>% 
  mutate(Occurrence = as.factor(Occurrence))

TEST <- lda_PA_dropper(abvIng, path = '../results/lda/', column_range = 7:22)

mapme <- left_join(select(abvi, ID, geometry), TEST)

ggplot() +
  geom_sf(data = mapme, aes(color = Occurrence))

table(TEST$Occurrence)

```