### 2024-01-10 scripts/predict_models.Rmd

---
title: "Predict Models"
author: "steppe"
date: "2024-01-02"
output: html_document
---

```{r}
library(tidyverse)
library(terra)
library(sf)
library(randomForest)
source('functions.R')

# unix::rlimit_as(1e32)
# setGDALconfig(c("GDAL_MAX_DATASET_POOL_SIZE"), c(""))
```

```{r read in predictors}
p <- '~/Documents/WesternPlantPredictors2'
paths <- list.dirs(p, recursive = T)
paths <- paths[2:length(paths)]
paths <- paths[! grepl('chelsa$|veg_cover$|soilgrids$|sand$|silt$|clay$|phh2o$|cfvo$|soc$', paths) ]

files <- lapply(paths, function(x){ file.path(x, list.files(x)) })
names(files) <- basename(paths)

layers <- lapply(files, vrt)
layers <- rast(layers)
names(layers) <- basename(paths)

rm(paths, files)
```

```{r read in models}

pmodels <- '../results/rf_models'
f <- file.path(pmodels, list.files(pmodels))
f <- data.frame(
  path = f,  
  species = gsub('-.*$', '', basename(f)), 
  time = gsub('[A-Za-z]', '', basename(f))
  ) |>
  mutate(time = gsub('_-|[.]', '', time),
         time = gsub('_', ' ', time), 
         time = as.POSIXct(time)) |>
  group_by(species) |>
  slice_max(order_by = time, n = 1) |>
  pull(path)

occpath <- '../data/raw/occurrence/combined_records'
occ_files <- list.files(occpath, pattern = 'shp$')

models_occ <- data.frame(
  path2model = f, occurrence_data = occ_files
) %>% 
  mutate(
    species = gsub('-.*$', '', basename(path2model)),
    match = if_else(species == gsub('[.]shp$', '', occurrence_data), T, F )
  ) # all look good!

models_occ <- split(models_occ, f = models_occ$species)

predict_wrapper <- function(x){

  sdm <- readRDS(x$path2model)
  occ_path <- file.path(occpath, x$occurrence_data)

  # read in occurrences and subset prediction to 100 miles from border
  bound <- sf::st_read(occ_path, quiet = TRUE) |>
    dplyr::filter(Occurrence == 1) |>
    sf::st_union() |>
    sf::st_convex_hull() |>
    sf::st_buffer(80467) |>
    sf::st_bbox()

  pout <- '~/Documents/suitability_maps'
  layer_subset <- terra::subset(layers, term_grab(sdm))
  layer_subset <- terra::crop(layer_subset, bound)

  OneClass <- \(...) predict(...)[,2]
  terra::predict(
    layer_subset, sdm, cpkgs="randomForest",
    type = 'prob', cores = 1, f = OneClass,
    filename = file.path(pout, paste0(x$species, '-', gsub(' ', '_', Sys.time()), '.tif')),
    wopt = c(names = 'predicted_suitability'))

}

lapply(models_occ[30:335], predict_wrapper)

```