### 2024-03-13 scripts/generate_absences.Rmd

---
title: "Generate Real and Pseudo-absences"
author: "steppe"
date: "2023-12-06"
output: html_document
---

```{r load packages}
library(tidyverse)
library(sf)
library(terra)
source('functions.R')
```

```{r Create simplified BLM surface management, eval = F}
p1 <- '/media/steppe/hdd/geospatial/SMA_WM.gdb'

st_layers(p1)
ag <- st_read(p1, 'SurfaceMgtAgy_BLM') %>% 
  filter(ADMIN_ST != 'AK') %>% 
  rename(geometry = SHAPE) %>% 
  select(geometry) %>% 
  st_make_valid() %>% 
  st_union()

ag <- st_simplify(ag, dTolerance = 1000)
plot(ag)

pout <- '/media/steppe/hdd/SDM_restorations/data/raw/coarse_blm/'
st_write(ag, dsn = paste0(pout, 'coarse_surface.shp'))
```

We use absences while creating species distribution models. 
Both true absences, acquired from vegetation monitoring plots, and pseudo absences generated via geographic and environmental space are used.

```{r load aim data for absences, eval = F}

p  <- '../data/raw/2021_AIM_Terrestrial/AIMTerrestrial9-1-22.gdb'
AIM_points <- st_read(dsn = p, layer = 'TerrADat', quiet = T) %>%
  select(PrimaryKey, DateVisited, geometry = Shape) %>% 
  st_transform(5070)

pout <- '/media/steppe/hdd/SDM_restorations/data/raw/coarse_blm/'
blm_surf <- st_read(paste0(pout, 'coarse_surface.shp'), quiet = TRUE)
blm_surf <- st_simplify(blm_surf, dTolerance = 500) # just make it a bit smaller. 
blm_surf <- st_make_valid(blm_surf)

rm(p, pout)
```

After we have removed spurious occurrence records (see chunk: 'subset occurrence records for final clean batch', in clean_records), we can generate absences for each species. 
```{r generate raw absences, eval = F}

# here we subset a terrestrial surface to our area of analysis - we don't want all of
# our random points to fall in the ocean

domain <- rast(nrows = 1, ncols = 1) # create a big empty raster, you can go in through sf too. 
ext(domain) <- c( -125.5, -100, 27,  50) # set the extent
crs(domain) <- "EPSG:4326" # define the projection
domain <- as.polygons(domain) |>  # convert to vector data
  st_as_sf() |> # convert to simple feature
  st_bbox()

land <- rnaturalearth::ne_download(type= 'land', scale = 10, category= 'physical') |>
  sf::st_as_sf()
land <- st_crop(land, domain) |>
  st_transform(5070)

rm(domain)

records <- st_read('../data/raw/occurrence/thinned/thinned_occurrences3.shp', quiet = TRUE)
records <- records %>% 
  group_by(taxon) %>% 
  filter(n() >= 15)
records <- split(records, f = records$taxon)

out1 <- lapply(records, absence_drawer, terrestrial = land,  bg_abs = 0.15, rand_dist = 1000)

out <- bind_rows(out1) %>% 
  rename(taxon = Taxon) %>% 
  mutate(date = as.Date('2023-01-09'), .before = geometry)

st_write(out, '../data/raw/occurrence/absences/absences1km3.shp', append = FALSE)

rm(out1, blm_surf, AIM_points, records)
```

```{r import raster stack, eval = F}

p <- '~/Documents/WesternPlantPredictors2'
paths <- file.path(p, list.files(p, recursive = F))
files <- lapply(paths, function(x){ file.path(x, list.files(x)) })
names(files) <- basename(paths)
bio <- lapply(files, vrt)
bio <- rast(bio)
rm(p, paths, files)

```

```{r read in predictors2 }
p <- '~/Documents/WesternPlantPredictors2'
paths <- list.dirs(p, recursive = T)
paths <- paths[2:length(paths)]
paths <- paths[! grepl('chelsa$|veg_cover$|soilgrids$|sand$|silt$|clay$|phh2o$|cfvo$|soc$', paths) ]

files <- lapply(paths, function(x){ file.path(x, list.files(x)) })
names(files) <- basename(paths)

layers <- lapply(files, vrt)
layers <- rast(layers)
names(layers) <- basename(paths)

rm(paths, files)
```

```{r subset rasterstack, eval = F}
vif_vals <- spatSample(bio, size = 5000, values = TRUE) # this steps takes forever... 
vif_vals <- vif_vals[ complete.cases(vif_vals), ] # 4148 records used
collinear_probs <- usdm::vifstep(vif_vals, th = 10)
bio <- usdm::exclude(bio, collinear_probs)

write_rds(collinear_probs, '../results/VIF_PA_drop.rda')

rm(p, paths, files)
```

```{r subset rasterstack 2, eval = F}
vif_vals <- spatSample(layers, size = 5000, values = TRUE) # this steps takes forever... 
vif_vals <- vif_vals[ complete.cases(vif_vals), ] # 4148 records used
collinear_probs <- usdm::vifstep(vif_vals, th = 10)
layers <- usdm::exclude(layers, collinear_probs)

write_rds(collinear_probs, '../results/VIF_PA_drop.rda')

rm(p, paths, files)
```

```{r import potential records}

collinear_probs <- read_rds('../results/VIF_PA_drop.rda')
layers <- usdm::exclude(layers, collinear_probs)

absences <- st_read('../data/raw/occurrence/absences/absences1km3.shp', quiet = TRUE)
presences <- st_read('../data/raw/occurrence/thinned/thinned_occurrences3.shp', quiet = TRUE) %>% 
  mutate(Occurrence = 1, 
         PtType = if_else(is.na(PrmryKy), 'AIM Presence', 'Opportunistic Presence'), .before = geometry)

recs <- bind_rows(absences, presences) %>% 
  mutate(taxon = str_replace(taxon, 'Dudleya cymosa ssp. cymosa', 'Dudleya cymosa')) %>% 
  group_by(taxon) %>% 
  mutate(ID = 1:n())

recs <- recs %>% 
  mutate(
  taxon = str_replace(taxon, 'Aslcepias subulata', 'Asclepias subulata'),
  taxon = str_replace(taxon, 'Elymus violaceus', 'Elymus trachycaulus'),
  taxon = str_replace(taxon, 'Encelia actonii$', 'Encelia actoni'),
  taxon = str_replace(taxon, 'Hillaria', 'Hilaria')
  ) %>% 
  filter(! taxon %in% c('Ephedra antisyphilitica', 'Ephedra fasciculata', 'Ephedra coryi', 'Ephedra cutleri',
                        'Ephedra funerea', 'Ephedra nevadensis', 'Ephedra torreyana', 'Ephedra californica',
                        'Ephedra trifurca', 'Ephedra trifurcata', 'Ephedra viridis', 'Ephedra aspera'))

set.seed(27)
cn <- names(layers)
splicies <- split(recs, f = recs$taxon)
names(splicies)
lapply(splicies[8:23], lda_PA_dropper, path = '../results/lda3/', col_names = cn)

rm(absences, presences, collinear_probs)
```

```{r remove pseudo absences in presence niche space, eval = F}

set.seed(27)
cn <- names(bio)
splicies <- split(recs, f = recs$taxon)

lapply(splicies, lda_PA_dropper, path = '../results/lda/', col_names = cn)
```