### 2024-04-03 scripts/GroundTruthSampling.Rmd
working oon 3 layer decisions

---
title: "Ground Truth Sampling Cells Colorado Plateau"
author: "steppe"
date: "2024-03-31"
output: html_document
---

To determine whether SDM's and hydrological basins can help find native plant populations field tests were conducted on the Northern Colorado Plateau.

```{r}
library(sf)
library(terra)
```

```{r import species data}
p <- '/media/steppe/hdd/2024SOS_CrewGeospatial/Plateau – Senior Botanist/Geodata'
f <- paste0(p, '/Species/SDM/', 
       list.files(paste0(p, '/Species/SDM/'), pattern = '.shp$'))
dat <- lapply(f, st_read, quiet = T) 
dat <- lapply(dat, '[', 2) 
dat <- lapply(dat, terra::vect)
names(dat) <- gsub('.shp$', '', basename(f))
# dat <- bind_rows(dat, .id = 'Taxon')

rm(f)
```

```{r import study extent and convert polygons to rasters}

blm_surf <- st_read(paste0(p, '/Admin/Surface/BLM_Surface.shp'), quiet = T) |>
  dplyr::filter(Unit_Nm %in% c('Grand Junction Field Office', 'Uncompahgre Field Office', 'Vernal Field Office', 'Price Field Office'))

r <- rast(blm_surf, resolution = 90, nlyrs = 3) # create 30 meter resolution raster
datr <- lapply(dat, rasterize, r, field = 'Rank')
datr <- rast(datr)
r[[1]] <- app(ifel(datr == 1, 1, NA), sum, na.rm = TRUE)
r[[2]] <- app(ifel(datr %in% c(2, 3, 4), 1, NA, na.rm = TRUE))
datr2 <-

rm(dat)
```