### 2024-04-03 scripts/GroundTruthSampling.Rmd
summary lyr 3

---
title: "Ground Truth Sampling Cells Colorado Plateau"
author: "steppe"
date: "2024-03-31"
output: html_document
---

To determine whether SDM's and hydrological basins can help find native plant populations field tests were conducted on the Northern Colorado Plateau.

```{r}
library(sf)
library(terra)
```

```{r import species data}
p <- '/media/steppe/hdd/2024SOS_CrewGeospatial/Plateau – Senior Botanist/Geodata'
f <- paste0(p, '/Species/SDM/', 
       list.files(paste0(p, '/Species/SDM/'), pattern = '.shp$'))
dat <- lapply(f, st_read, quiet = T) 
dat <- lapply(dat, '[', 2) 
dat <- lapply(dat, terra::vect)
names(dat) <- gsub('.shp$', '', basename(f))
# dat <- bind_rows(dat, .id = 'Taxon')

rm(f)
```

```{r import study extent and convert polygons to rasters}

blm_surf <- st_read(paste0(p, '/Admin/Surface/BLM_Surface.shp'), quiet = T) |>
  dplyr::filter(Unit_Nm %in% c('Grand Junction Field Office', 'Uncompahgre Field Office', 'Vernal Field Office', 'Price Field Office'))

r <- rast(blm_surf, resolution = 90, nlyrs = 3) # create 30 meter resolution raster
names(r) <- c('SumOccupiedPatches', 'Sum1stOrderNeighboringPatches', 'CommonessInversion')
datr <- lapply(dat, rasterize, r, field = 'Rank')
datr <- rast(datr)

# count the occupied patches by overlap  
r[[1]] <- app(ifel(datr == 1, 1, NA), sum, na.rm = TRUE)

# count the patches which are first order neighbors from a known-populated locality  
r[[2]] <- app(ifel(datr %in% c(2, 3, 4), 1, NA), sum, na.rm = TRUE)

# How common are occupied + 1st order neighbors for each species? i.e. which are less common
# and should be considered more importantly during scouting ??

datr <- ifel(datr <= 4, 1, NA)
tab <- freq(datr)

tab$InverseWt <- 1 - ( tab$count / max(tab$count))
tab$NegativeInverseWt <- round( 1 / (max(tab$InverseWt+1) - tab$InverseWt), 2) * 100
# extreme skew rescale with 'negative transform', keep digits few for memory.

par(mfrow=c(1,2)) 
hist(tab$InverseWt, main = 'Raw', xlab = 'Weight') 
hist(tab$NegativeInverseWt, main = 'Negative Transform', xlab = 'Weight')

for (i in seq_along(1:dim(datr)[3])){ 
  datr[[i]] <- ifel(datr[[i]] == 1, tab$NegativeInverseWt[i], NA) 
}

r[[3]] <- app(datr, sum, na.rm = TRUE)
names(r) <- c('SumOccupiedPatches', 'Sum1stOrderNeighboringPatches', 'SumInverseCommonness')
plot(r)

writeRaster(r, '../results/Summary_Rasters/COPL.tif')

rm(datr, tab, i)
```