### 2024-09-13 scripts/Results.Rmd
conceptual figs

---
title: "Results"
author: "steppe"
date: "2024-08-12"
output: pdf_document
---

```{r load libraries}
library(tidyverse)
library(sf)
library(terra)
```

## Import and prepare data for analysis.

```{r Subset to relevant species}
scouting <- read.csv('../data/SOS/Scouting.csv') |>
  mutate(taxa = gsub(' ', '_', taxa)) |>
  select(c(1:10, 22, 23)) |>
  filter(taxa != 'Achnatherum_hymenoides')

f <- list.files('../results/suitability_maps')
f <- f[ str_detect(f, '1k')]

f <- data.frame(
  f_suitability = paste0('../results/suitability_maps/', f), 
  species = gsub('1k-.*', '', f)
)

scouting <- filter(scouting, taxa %in% f$species) |>
  sf::st_as_sf(coords = c('LONGITUDE_DECIMAL', 'LATITUDE_DECIMAL'), crs = 4326) |>
  st_transform(5070)

# here are the ranked patches
f_rp <- list.files('../results/rankedPatches', pattern = '.shp')
f_rp <- data.frame(
  f_rankedp = paste0('../results/rankedPatches/', f_rp), 
  species = gsub('[.]shp$', '', f_rp)
)

f_pm <- list.files('../results/patch_metrics')
f_pm <- data.frame(
  f_patchMetrics = paste0('../results/patch_metrics/', f_pm[grepl('patch', f_pm)]), 
  f_classMetrics = paste0('../results/patch_metrics/', f_pm[grepl('class', f_pm)]) 
) |>
  mutate(species = gsub('-.*$', '', basename(f_patchMetrics)))

files <- left_join(f, f_rp, by = 'species') |>
  left_join(f_pm, by = 'species') |>
  relocate(species, .before = 1) |>
  filter(species %in% scouting$taxa)

rm(f, f_rp, f_pm)
```

```{r Extract Values to Results}

scouting <- left_join(scouting, files, by = c('taxa' = 'species')) |>
  mutate(Presence = if_else(POPULATION_SIZE > 0, 1, 0), .after = POPULATION_SIZE) |>

  # penstemon palmeri are all seeded outside their native range and 'helped' along
  # during establishment we will remove them from the analysis

  filter(taxa != 'Penstemon_palmeri')

scout_split <- split(scouting, scouting$taxa)
# we are going to discard species with fewer than 3 observations. 
scout_split <- scout_split[lapply(scout_split, nrow) >= 3]

#' return predicted suitability scores and patch ID's for scouting points
#' 
#' Use this function to return some modeled values from scouting points observations
#' @param x an sf/tibble/data frame split by species, and including columns with 
#' paths to required files. 
extractR <- function(x){

  suitability <- terra::rast(sf::st_drop_geometry(x$f_suitability[1]))
  patch_ranks <- sf::st_read(sf::st_drop_geometry(x$f_rankedp[1]), quiet = T)

  x <- sf::st_join(x, patch_ranks)
  suitability <- terra::extract(suitability, x)$predicted_suitability

  # now we can join the class and patch metrics to each population 
  # which they were calculated for. 
  IDs <- tidyr::drop_na(x, PchIDComb) |>
    dplyr::pull(PchIDComb)

  patch <- read.csv(sf::st_drop_geometry(x$f_patchMetrics[1]) ) |>
    dplyr::filter(id %in% IDs) |>
    tidyr::pivot_wider(values_from = value, names_from = metric)
  class <- read.csv(sf::st_drop_geometry(x$f_classMetrics[1])) |>
    tidyr::pivot_wider(names_from = metric, values_from = value)

  if(nrow(patch) == 0){
    patch <- data.frame(id = NA, cai = NA, enn = NA, frac = NA, para = NA )
  }

  lsm <- cbind(patch, class) |>
    dplyr::mutate(id = as.character(id))

  wide <- dplyr::mutate(
    x, 
    suitability = suitability, 
    .before = geometry) |>
    dplyr::select(-f_suitability, -f_rankedp, -f_patchMetrics, -f_classMetrics) |>
    dplyr::left_join(lsm, by = c('PchIDComb' = 'id' ))

  return(wide)

}

scouting <- lapply(scout_split, extractR) 
scouting <- dplyr::bind_rows(scouting)

```

```{r write out some species for further data acquisition, eval = F}
sort( gsub('_', ' ', names(scout_split))) # these will be the species we make notes on.

data.frame(
  Species = sort( gsub('_', ' ', names(scout_split))), 
  'Pres-Abs' = '',
  Census = ''
)|>
  write_csv('../data/SOS/Absence_sampling.csv')

```

```{r determine a null model for each species, eval = F}
#' return a large sample from a raster to plot the amount of suitable habitat 
#' 
#' sample many rasters to make a plot showing the cumulative amount of suitable habitat
#' @param x a data frame split by species including columns with paths to files. 
prcntSuitable <- function(x){

  r <- terra::rast(sf::st_drop_geometry(x$f_suitability[1]))
  Suitability = spatSample(r, size = 50000, na.rm = TRUE)
  q <- quantile(Suitability$predicted_suitability, probs = seq(0.01, 1, 0.01))

  sample <- 
    data.frame(
      Species = sf::st_drop_geometry(x$taxa[1]), 
      Percent = 1:100, 
      Records = as.numeric(q)
    )

  return(sample)

}

suitability_amounts <- lapply(scout_split, prcntSuitable)
saveRDS(suitability_amounts, file = '../data/processed/prcntSDMsurfaceSuitable.rds')
```

```{r}

habitat_specialist <- data.frame(
  suitability = c(
    seq(from= 0, to  = 0.1, length.out = 70), 
    seq(from= 0.1, to  = 0.4, length.out = 15), 
    seq(from= 0.4, to  = 0.8, length.out = 10),
    seq(from= 0.8, to  = 1, length.out = 5)
  ),
  percent = 1:100
)

habitat_generalist <- data.frame(
  suitability = c(
    seq(from= 0, to  = 0.5, length.out = 40), 
    seq(from= 0.6, to  = 0.7, length.out = 30), 
    seq(from= 0.7, to  = 0.9, length.out = 20),
    seq(from= 0.9, to  = 1, length.out = 10)
  ),
  percent = 1:100
)

ggplot() + 
  geom_smooth(data = habitat_specialist, aes(x = percent, y = suitability)) + 
  geom_smooth(data = habitat_generalist, aes(x = percent, y = suitability)) +
  theme(
    plot.title = element_text(hjust = 0.5),
    aspect.ratio = 1/1)

```

```{r}

suitability_amounts <- readRDS('../data/processed/prcntSDMsurfaceSuitable.rds')
suitability_amounts <- bind_rows(suitability_amounts)
remotes::install_github("bcgov/elucidate")

y1 <- rnorm(1:10, 100, 10)
elucidate::mean_ci(y1)

recs <- filter(suitability_amounts, Percent == 50) |>
  pull(Records)

as.numeric( elucidate::mean_ci(recs, replicates = 1000, ci_level = 0.99) )

prctn <- split(suitability_amounts, f = suitability_amounts$Percent)
suitability_means <- lapply(prctn, \(x)
  data.frame(as.list(
    elucidate::mean_ci(x$Records, replicates = 1000, ci_level = 0.99))
  )
) |>
  bind_rows() |>
  mutate(Percent = 1:100)

fp <- scouting[scouting$futurePotential=='Yes',] 
fp <- arrange(fp, suitability) |>
  st_drop_geometry() |>
  select(suitability) |>
  mutate(obs = 1:n() / n())

linecols <- c(
  'Species' = 'grey50',
  "Mean" = "#093824", 
  "95% CI" = "#73BA9B",
  "Populations" = "#FF2C55"
  )

suitability_means <- suitability_means |>
  mutate(Null_mod = mean*Percent)

ggplot(suitability_amounts,
       aes(x = Percent, y = Records, color = Species, group = Species)) +
  geom_line(lty = 1) +
  scale_color_grey(guide = 'none') + 
  geom_ribbon(data = suitability_means, aes(ymin = lower, ymax = upper, x = Percent),
              fill = '#73BA9B', inherit.aes = F, alpha = 0.5) + 
  geom_line(data = suitability_means, 
            aes(x = Percent, y =  mean, color = 'a'), 
            color = '#093824', inherit.aes = F, lwd = 1) + 
  geom_line(data = suitability_means, aes(x = Percent, y = Null_mod/100),
            color = 'black', inherit.aes = F,  lwd = 1) +
  geom_line(data = fp, aes(x = suitability*100, y = obs),
            color = '#FF2C55', inherit.aes = F,  lwd = 1) +

  labs(
    title = 'Populations are found in\nareas with higher suitability',
    x = "Raster Cells (cumulative)",
    y = 'Predicted Suitability (cumulative)') + 
  theme_classic() +
  scale_y_continuous(breaks = seq(0, 1, by = 0.2), sec.axis = sec_axis(~.*100,
    name = 'Populations Found (cumulative)',
    breaks = seq(0, 100, by = 20), 
    labels = \(x) paste0(x, "%"))
  )  +

  scale_x_continuous(breaks = seq(0, 100, by = 20), 
                     labels = \(x) paste0(x, "%")) + 
  theme(
    plot.title = element_text(hjust = 0.5),
    aspect.ratio = 1/1)

rm(linecols)
```

```{r}
suitability_means[ which.min(abs(suitability_means$Percent-50)), 'mean'] # on average 50% of cells have lower probabilities of being suitable than 0.23

suitability_means[ which.min(abs(suitability_means$Percent-80)), 'mean'] # on average 80% of cells have lower probabilities of being suitable than 0.614

fp[ which.min(abs(fp$obs-0.2)), ] # 80% of all populations are found in areas with suitability > 65.2%

fp[ which.min(abs(fp$obs-0.5)), ] # 50% of all populations are found in areas with suitability > 83.6%

fp[ which.min(abs(fp$obs-0.8)), ] # the final 20 % of populations are found in areas with suitability > 94.6%
```

## Determine which points were used in training the models

These points will need to be removed because they are already known to be presences...

## Analysis

1) What is the relationship between habitat suitability and presence of species?

```{r}

fp <- scouting[scouting$futurePotential=='Yes',] 
quants <- quantile(fp$suitability, na.rm = TRUE, probs = c(0.25, 0.5, 0.75))

hist(fp$suitability)
abline(v = quants[[1]], lwd = 2)
abline(v = quants[[2]], lwd = 2)

```

```{r}

```

2) What is the relationship between Patch Rank and presence of species?

- can we test moving the distance patches into their own categories ??? 
```{r}
hist(fp$Rank)
```

3) What is the relationship between Patch metrics and presence of species?

```{r}

```

The second main part relates to population size

4) What is the relationship between habitat suitability and population size? 
```{r}
scout1 <- mutate(
  scout1, 
  taxa = as.factor(taxa), 
  COLL_ID = as.factor(COLL_ID))

plot(scout1$suitability, scout1$POPULATION_SIZE)

m <- lm(POPULATION_SIZE ~ suitability + taxa, data = scout1)
summary(m)

pois.mod <- lme4::glmer(POPULATION_SIZE ~ suitability + (1 | taxa), data = scout1, family = poisson)
summary(pois.mod)
anova(pois.mod)

pois.mod_int <- lme4::glmer(POPULATION_SIZE ~ suitability + (1 + suitability|taxa), data = scout1, family = poisson)
summary(pois.mod_int)

plot(pois.mod_int)
plot(residuals(pois.mod_int))

qqnorm(resid(pois.mod_int))# QQ-plot
qqline(resid(pois.mod_int))

AIC(pois.mod, pois.mod_int)
anova(pois.mod, pois.mod_int)
```

4) What is the relationship between patch metrics and population size? 
```{r}

scout_long <- pivot_longer(scouting, cai:enn_mn)

ggplot(scout_long, aes(y = POPULATION_SIZE, x = value)) +
  geom_point() +
  facet_grid(. ~ name)

```