### 2024-09-13 scripts/Results.Rmd
crazy null plot

---
title: "Results"
author: "steppe"
date: "2024-08-12"
output: pdf_document
---

```{r load libraries}
library(tidyverse)
library(sf)
library(terra)
```

## Import and prepare data for analysis.

```{r Subset to relevant species}
scouting <- read.csv('../data/SOS/Scouting.csv') |>
  mutate(taxa = gsub(' ', '_', taxa)) |>
  select(c(1:10, 22, 23)) |>
  filter(taxa != 'Achnatherum_hymenoides')

f <- list.files('../results/suitability_maps')
f <- f[ str_detect(f, '1k')]

f <- data.frame(
  f_suitability = paste0('../results/suitability_maps/', f), 
  species = gsub('1k-.*', '', f)
)

scouting <- filter(scouting, taxa %in% f$species) |>
  sf::st_as_sf(coords = c('LONGITUDE_DECIMAL', 'LATITUDE_DECIMAL'), crs = 4326) |>
  st_transform(5070)

# here are the ranked patches
f_rp <- list.files('../results/rankedPatches', pattern = '.shp')
f_rp <- data.frame(
  f_rankedp = paste0('../results/rankedPatches/', f_rp), 
  species = gsub('[.]shp$', '', f_rp)
)

f_pm <- list.files('../results/patch_metrics')
f_pm <- data.frame(
  f_patchMetrics = paste0('../results/patch_metrics/', f_pm[grepl('patch', f_pm)]), 
  f_classMetrics = paste0('../results/patch_metrics/', f_pm[grepl('class', f_pm)]) 
) |>
  mutate(species = gsub('-.*$', '', basename(f_patchMetrics)))

files <- left_join(f, f_rp, by = 'species') |>
  left_join(f_pm, by = 'species') |>
  relocate(species, .before = 1) |>
  filter(species %in% scouting$taxa)

rm(f, f_rp, f_pm)
```

```{r Extract Values to Results}

scouting <- left_join(scouting, files, by = c('taxa' = 'species')) |>
  mutate(Presence = if_else(POPULATION_SIZE > 0, 1, 0), .after = POPULATION_SIZE) |>

  # penstemon palmeri are all seeded outside their native range and 'helped' along
  # during establishment we will remove them from the analysis

  filter(taxa != 'Penstemon_palmeri')

scout_split <- split(scouting, scouting$taxa)
# we are going to discard species with fewer than 3 observations. 
scout_split <- scout_split[lapply(scout_split, nrow) >= 3]

#' return predicted suitability scores and patch ID's for scouting points
#' 
#' Use this function to return some modeled values from scouting points observations
#' @param x an sf/tibble/data frame split by species, and including columns with 
#' paths to required files. 
extractR <- function(x){

  suitability <- terra::rast(sf::st_drop_geometry(x$f_suitability[1]))
  patch_ranks <- sf::st_read(sf::st_drop_geometry(x$f_rankedp[1]), quiet = T)

  x <- sf::st_join(x, patch_ranks)
  suitability <- terra::extract(suitability, x)$predicted_suitability

  # now we can join the class and patch metrics to each population 
  # which they were calculated for. 
  IDs <- tidyr::drop_na(x, PchIDComb) |>
    dplyr::pull(PchIDComb)

  patch <- read.csv(sf::st_drop_geometry(x$f_patchMetrics[1]) ) |>
    dplyr::filter(id %in% IDs) |>
    tidyr::pivot_wider(values_from = value, names_from = metric)
  class <- read.csv(sf::st_drop_geometry(x$f_classMetrics[1])) |>
    tidyr::pivot_wider(names_from = metric, values_from = value)

  if(nrow(patch) == 0){
    patch <- data.frame(id = NA, cai = NA, enn = NA, frac = NA, para = NA )
  }

  lsm <- cbind(patch, class) |>
    dplyr::mutate(id = as.character(id))

  wide <- dplyr::mutate(
    x, 
    suitability = suitability, 
    .before = geometry) |>
    dplyr::select(-f_suitability, -f_rankedp, -f_patchMetrics, -f_classMetrics) |>
    dplyr::left_join(lsm, by = c('PchIDComb' = 'id' ))

  return(wide)

}

scouting <- lapply(scout_split, extractR) 
scouting <- dplyr::bind_rows(scouting)

```

```{r write out some species for further data acquisition, eval = F}
sort( gsub('_', ' ', names(scout_split))) # these will be the species we make notes on.

data.frame(
  Species = sort( gsub('_', ' ', names(scout_split))), 
  'Pres-Abs' = '',
  Census = ''
)|>
  write_csv('../data/SOS/Absence_sampling.csv')

```

```{r determine a null model for each species, eval = F}
#' return a large sample from a raster to plot the amount of suitable habitat 
#' 
#' sample many rasters to make a plot showing the cumulative amount of suitable habitat
#' @param x a data frame split by species including columns with paths to files. 
prcntSuitable <- function(x){

  r <- terra::rast(sf::st_drop_geometry(x$f_suitability[1]))
  Suitability = spatSample(r, size = 50000, na.rm = TRUE)
  q <- quantile(Suitability$predicted_suitability, probs = seq(0.01, 1, 0.01))

  sample <- 
    data.frame(
      Species = sf::st_drop_geometry(x$taxa[1]), 
      Percent = 1:100, 
      Records = as.numeric(q)
    )

  return(sample)

}

suitability_amounts <- lapply(scout_split, prcntSuitable)
saveRDS(suitability_amounts, file = '../data/processed/prcntSDMsurfaceSuitable.rds')
```

```{r}

suitability_amounts <- readRDS('../data/processed/prcntSDMsurfaceSuitable.rds')
suitability_amounts <- bind_rows(suitability_amounts)

prctn <- split(suitability_amounts, f = suitability_amounts$Percent)
suitability_means <- lapply(prctn, \(x)
  data.frame(as.list(
    elucidate::mean_ci(x$Records, replicates = 1000, ci_level = 0.99))
  ) 
) |>
  bind_rows() |>
  mutate(Percent = 1:100)

#################################################################################
fp <- scouting[scouting$futurePotential=='Yes',] 
fp <- arrange(fp, suitability) |>
  st_drop_geometry() |>
  select(suitability) |>
  mutate(obs = 1:n() / n())

### Here we prep some display text which will help users interpret the plot ####
################################################################################
sm50 <- suitability_means[ which.min(abs(suitability_means$Percent-50)), ] 
# on average 50% of cells have lower probabilities of being suitable than 0.23

sm80 <- suitability_means[ which.min(abs(suitability_means$Percent-80)), ] 
# on average 80% of cells have lower probabilities of being suitable than 0.614

fp_2 <- fp[ which.min(abs(fp$obs-0.2)), ] # 80% of all populations are found in areas with suitability > 65.2%
fp_5 <- fp[ which.min(abs(fp$obs-0.5)), ] # 50% of all populations are found in areas with suitability > 83.6% 
fp_8 <- fp[ which.min(abs(fp$obs-0.8)), ] # the final 20 % of populations are found in areas with suitability > 94.6%

labs <- data.frame(
  x = c(
    round(fp_2$obs, 2)*100 -10, 
    round(fp_5$obs, 2)*100 - 10, 
    sm50$Percent+15, 
    sm80$Percent+15),
  y = c(
    round(fp_2$suitability, 2) + 0.1,
    round(fp_5$suitability, 2) + 0.1, 
    sm50$mean - 0.15, 
    sm80$mean - 0.15),
  label = c(
    paste0(100- (round(fp_2$obs, 2)*100), '% of pops.\nwere found in areas\nwith suitability > ', round(fp_2$suitability, 2)),  
    paste0(round(fp_5$obs, 2)*100, '% of pops.\nwere found in areas\nwith suitability > ', round(fp_5$suitability, 2)),
    paste0(sm50$Percent, '% of cells have\nprobabilities of being\nsuitable lower than ', round(sm50$mean, 2)), 
    paste0(sm80$Percent, '% of cells have\nprobabilities of being\nsuitable lower than ', round(sm80$mean, 2))
  )
)

################################################################################

# Create the mean suitability values across all taxa, this will illustrate the
# distribution of modeled suitable habitat across 'common species' in the West based
# on our input data and modelling approach. 
suitability_means <- suitability_means |>
  mutate(Null_mod = mean*Percent)

linecols <- c( # define our palette
  'Species' = 'grey50',
  "Mean" = "#093824", 
  "95% CI" = "#73BA9B",
  "Populations" = "#FF2C55"
  )

ggplot(suitability_amounts,
       aes(x = Percent, y = Records, color = Species, group = Species)) +
  geom_line(lty = 1) +
  scale_color_grey(guide = 'none') + 
  geom_ribbon(data = suitability_means, aes(ymin = lower, ymax = upper, x = Percent),
              fill = '#73BA9B', inherit.aes = F, alpha = 0.5) + 
  geom_line(data = suitability_means, 
            aes(x = Percent, y =  mean, color = 'a'), 
            color = '#093824', inherit.aes = F, lwd = 1) + 
  geom_line(data = suitability_means, aes(y = Percent/100, x = Null_mod),
            color = '#d95f02', inherit.aes = F,  lwd = 1, lty = 53, linemitre = 2) +
  geom_line(data = fp, aes(y = suitability, x = obs*100),
            color = '#7570b3', inherit.aes = F,  lwd = 1, lty = 53) +

  labs(
    title = 'Populations are found in\nareas with higher suitability',
    x = "Raster Cells (cumulative)",
    y = 'Predicted Suitability') + 
  theme_classic() +
  scale_y_continuous(
    breaks = seq(0, 1, by = 0.2)
  )  +

  scale_x_continuous(
    breaks = seq(0, 100, by = 20), labels = \(x) paste0(x, "%"), 
          sec.axis = sec_axis(~., name = 'Populations Found (cumulative)', 
                        breaks = seq(0, 100, by = 20), labels = \(x) paste0(x, "%"))) + 
  theme(
    plot.title = element_text(hjust = 0.5),
    aspect.ratio = 1/1) +

  geom_segment(
    aes(
      xend = fp_2$obs*100, 
      yend = fp_2$suitability, 
      x = round(fp_2$obs, 2)*100 - 10, 
      y = fp_2$suitability + 0.1),
                  arrow = arrow(length = unit(0.25, "cm")), color = 'black') + 
  geom_segment(
    aes(
      xend = fp_5$obs*100, 
      yend = fp_5$suitability, 
      x = round(fp_5$obs, 2)*100 - 10, 
      y = fp_5$suitability + 0.1),
    arrow = arrow(length = unit(0.25, "cm")), color = 'black') +

  geom_segment(aes(xend = sm50$Percent, yend = sm50$mean, 
                   x = sm50$Percent+15, y = sm50$mean - 0.15),
                  arrow = arrow(length = unit(0.25, "cm")), color = 'black') +

  geom_segment(aes(xend = sm80$Percent, yend = sm80$mean, 
                   x = sm80$Percent+15, y = sm80$mean - 0.15),
               arrow = arrow(length = unit(0.25, "cm")), color = 'black') +

  geom_label(data = labs, aes(x = x, y = y, label = label), inherit.aes = F, size = 2.5)

# rm(fp, fp_2, fp_5, fp_8, labs, prctn, sm50, sm80, linecols)
```

Raster cells (X axis, lower), represent units of the entire landscape, and the landscape predicts the amount of suitable habitat (Y axis) available for a species. 
Across all species, the amount of suitable habitat appears low, on average (green line) for any one species, 80% of the landscape has a less than 0.61 chance of being suitable. 
So on average only 20% of the landscape has high odds of being suitable habitat for an average species. 
We can multiple the probability of each raster cell having suitable habitat to generate an expectation of how many populations crews would find based on our models alone (orange line).
Using empirical data (purple line) we show that crews mostly found populations in areas with high predicted suitability, 80% of populations were in areas with suitability > 0.65, 50% in areas with modeled suitability > 0.84$. 
Where the purple line is above the orange line, it shows that crews found fewer populations than would be expected by our model; I wager most of this deviation is due to them simply not recording populations - a sampling artifact.

## Determine which points were used in training the models

These points will need to be removed because they are already known to be presences...

## Analysis

1) What is the relationship between habitat suitability and presence of species?

```{r}

fp <- scouting[scouting$futurePotential=='Yes',] 
quants <- quantile(fp$suitability, na.rm = TRUE, probs = c(0.25, 0.5, 0.75))

hist(fp$suitability)
abline(v = quants[[1]], lwd = 2)
abline(v = quants[[2]], lwd = 2)

```

```{r}

```

2) What is the relationship between Patch Rank and presence of species?

- can we test moving the distance patches into their own categories ??? 
```{r}
hist(fp$Rank)
```

3) What is the relationship between Patch metrics and presence of species?

```{r}

```

The second main part relates to population size

4) What is the relationship between habitat suitability and population size? 
```{r}
scout1 <- mutate(
  scout1, 
  taxa = as.factor(taxa), 
  COLL_ID = as.factor(COLL_ID))

plot(scout1$suitability, scout1$POPULATION_SIZE)

m <- lm(POPULATION_SIZE ~ suitability + taxa, data = scout1)
summary(m)

pois.mod <- lme4::glmer(POPULATION_SIZE ~ suitability + (1 | taxa), data = scout1, family = poisson)
summary(pois.mod)
anova(pois.mod)

pois.mod_int <- lme4::glmer(POPULATION_SIZE ~ suitability + (1 + suitability|taxa), data = scout1, family = poisson)
summary(pois.mod_int)

plot(pois.mod_int)
plot(residuals(pois.mod_int))

qqnorm(resid(pois.mod_int))# QQ-plot
qqline(resid(pois.mod_int))

AIC(pois.mod, pois.mod_int)
anova(pois.mod, pois.mod_int)
```

4) What is the relationship between patch metrics and population size? 
```{r}

scout_long <- pivot_longer(scouting, cai:enn_mn)

ggplot(scout_long, aes(y = POPULATION_SIZE, x = value)) +
  geom_point() +
  facet_grid(. ~ name)

```